---
title: "simulate incidence - check"
author: "Pierre Nouvellet"
date: "2023-09"
output:
  html_document:
    self_contained: no
---

```{r , include = FALSE, message = FALSE, warning = FALSE, error = FALSE}
rm(list=ls())


library(Hmisc)
library(EpiEstim)
library(incidence)
library(projections)
library(readxl)

remove.packages('MCMCEpiEstim')
devtools::install_github("pnouvellet/MCMCEpiEstim",build = TRUE)
library(MCMCEpiEstim)

```


# Simulated input data

## config for simulation

```{r}

# Scenarios <- as.data.frame(readxl::read_xlsx('../overdispersion_scenarios.xlsx',
#                                sheet = 1, range = 'A1:I60'))
Scenarios <- as.data.frame(readxl::read_xlsx('../overdispersion-epidemic-data/overdispersion_scenarios.xlsx',
                               sheet = 1, range = 'A1:I60'))

input <- Scenarios[1,]

si <- c(0,1)

```

## simulate Rt and I

```{r}

I0 <- incidence::as.incidence(x = 10, dates = 1, interval = 1)

```


# Poisson simulation

```{r}
Rt <- MCMCEpiEstim::Rt_linear(inp = input)
res <- MCMCEpiEstim::project_fct(I0 = I0, Rt = Rt, 
                                  n_loc = 1e2,#input$n_locations,
                                  t_max = nrow(Rt), si = si,p = 1,
                                  model = 'poisson')#,over = input$k)
I_NB <- res$I_true
```

check means

```{r}

plot(1:(nrow(I_NB)-1),apply(I_NB[2:(nrow(I_NB)),-1]/I_NB[1:(nrow(I_NB)-1),-1],1,mean,na.rm=TRUE))
lines(Rt$t,Rt$Rt)


```

check variances

```{r}

c(Rt$Rt[1]*I0$counts,sd(I_NB[2,-1])^2)
E_NB <- I_NB[1:(nrow(I_NB)-1),-1] * matrix(Rt$Rt,nrow = nrow(Rt),ncol = ncol(I_NB)-1,byrow = FALSE)
E_NB <- as.matrix(rbind(NA,E_NB))

Var_I <- data.frame(I = 1:50, Var = NA, n = NA)
temp <- as.matrix(I_NB)
temp[,1]<-NA
for(i in 1:nrow(Var_I)){
  f <- which(E_NB == Var_I$I[i])
  Var_I$Var[i] <- sd(temp[f])
  Var_I$n[i] <- length(f)
}

plot(Var_I$I,Var_I$Var, xlim = c(0,max(c(Var_I$I,Var_I$Var), na.rm=TRUE )),ylim = c(0,max(c(Var_I$I,Var_I$Var), na.rm=TRUE )) )
abline(a = 0, b = 1,col = 'red')

plot(Var_I$I,Var_I$n)


```

check residuals vs. Poisson

```{r}

layout(matrix(1:2, nrow = 1))
P1 <- diag_plot(I_NB, logged = TRUE, max_x=1, dist = 'poisson', Rt = Rt)
P2 <- diag_plot(I_NB, logged = FALSE, max_x=.1, dist = 'poisson', Rt = Rt)

P1$res_table  
P2$res_table  

P1$check$v1 <- P1$check$residual^2/P1$check$Exp
boxplot(P1$check$v1~P1$check$Rt)
m1<-lm(P1$check$v1~P1$check$Rt)
summary(m1)
```




# NB simulation

```{r}
k=.1
res <- project_fct(I0 = I0, Rt = Rt, 
                                  n_loc = 1e3,#input$n_locations,
                                  t_max = nrow(Rt), si = si,p = 1,
                                  model = 'negbin',over = k)
I_NB <- res$I_true
```

check means

```{r}

plot(1:(nrow(I_NB)-1),apply(I_NB[2:(nrow(I_NB)),-1]/I_NB[1:(nrow(I_NB)-1),-1],1,mean,na.rm=TRUE))
lines(Rt$t,Rt$Rt)


```

check variances

```{r}

c(Rt$Rt[1]*I0$counts,sd(I_NB[2,-1])^2)
E_NB <- I_NB[1:(nrow(I_NB)-1),-1] * matrix(Rt$Rt,nrow = nrow(Rt),ncol = ncol(I_NB)-1,byrow = FALSE)
E_NB <- as.matrix(rbind(NA,E_NB))

Var_I <- data.frame(I = 1:50, Var = NA, n = NA)
temp <- as.matrix(I_NB)
temp[,1]<-NA
for(i in 1:nrow(Var_I)){
  f <- which(E_NB == Var_I$I[i])
  Var_I$Var[i] <- sd(temp[f])
  Var_I$n[i] <- length(f)
}

plot(Var_I$I,Var_I$Var, xlim = c(0,max(c(Var_I$I,Var_I$Var), na.rm=TRUE )),ylim = c(0,max(c(Var_I$I,Var_I$Var), na.rm=TRUE )) )
x <- seq(0,max(c(Var_I$I,Var_I$Var), na.rm=TRUE ))
y <- 1*x*(1+1/k) # assume R of 1
abline(a = 0, b = 1,col = 'red')
lines(x,y,col = 'blue')

plot(Var_I$I,Var_I$n)


```

check residuals vs. Poisson


```{r}

layout(matrix(1:2, nrow = 1))
P1 <- diag_plot(I_NB, logged = TRUE, max_x=1, dist = 'poisson', Rt = Rt)
P2 <- diag_plot(I_NB, logged = FALSE, max_x=.1, dist = 'poisson', Rt = Rt)

layout(matrix(1:2, nrow = 1))
NB1 <- diag_plot(I_NB, logged = TRUE, max_x=1, dist = 'nb', Rt = Rt,k = k)
NB2 <- diag_plot(I_NB, logged = FALSE, max_x=.1, dist = 'nb', Rt = Rt, k = k)

P1
P2
NB1
NB2


NB1$check$v1 <- NB1$check$residual^2/NB1$check$Exp

boxplot(NB1$check$v1~NB1$check$Rt)
m1<-lm(NB1$check$v1~NB1$check$Rt)
summary(m1)

P1$check$v1 <- P1$check$residual^2/P1$check$Exp

boxplot(P1$check$v1~P1$check$Rt)
m1<-lm(P1$check$v1~P1$check$Rt)
summary(m1)

```







# check MCMC with NB simulation


# simulate

```{r}

Scenarios <- as.data.frame(readxl::read_xlsx('../overdispersion-epidemic-data/overdispersion_scenarios.xlsx',
                               sheet = 1, range = 'A1:I60'))
si <- c(0,1)

i=57

# print(i)
input <- Scenarios[i,]

input$R_min <- 0.65
input$k <- .1

I0 <- incidence::as.incidence(x = input$I0, dates = 1, interval = 1)

Rt <- MCMCEpiEstim::Rt_linear(inp = input)

res <- MCMCEpiEstim::project_fct(I0 = I0, Rt = Rt, 
                                 n_loc = input$n_locations,
                                 t_max = nrow(Rt), si = si,p = input$pi,
                                 model = 'negbin',over = input$k)
sim <- list(input = input, I0 = I0, Rt = Rt, res = res)


matplot((sim$res$I_true[,-1]))
matplot((sim$res$I_obs[,-1]))
input
  
```



```{r}


# parameters
# for EpiEstim
t_window <- 7
mean_prior <- 5
std_prior <- 5

res_EpiEstim <- fct_EpiEstim(I0 = sim$I0 , I = sim$res$I_obs, #I_true ,
                             t_window = t_window, 
                             mean_prior = mean_prior, std_prior = std_prior,
                             si = si, overlap = FALSE)


plot(Rt$t,Rt$Rt,type = 'l', ylim = c(0,2.5))

for(i in 1:sim$input$n_locations){
  
  Hmisc::errbar(x = res_EpiEstim[[i]]$t_start, 
                y = res_EpiEstim[[i]]$`Median(R)`,
                yplus = res_EpiEstim[[i]]$`Quantile.0.975(R)`,
                yminus = res_EpiEstim[[i]]$`Quantile.0.025(R)`,
                col = i, errbar.col = i, 
                add=TRUE)
}



```
diagnoistics for epiEstim

```{r}

I_NB <- as.matrix(sim$res$I_obs) #I_true)

Rts_epiestim <- matrix(unlist(lapply(res_EpiEstim, "[", ,'Mean(R)')), 
                       nrow = t_max, ncol = n_location, byrow = FALSE)

Rts_epiestim <- Rts_epiestim[seq(1,nrow(Rts_epiestim),by = 7),]
Rts_epiestim <- Rts_epiestim[rep(seq(2,nrow(Rts_epiestim)),each = 7),]

Rts_epiestim <- data.frame(t = 1:nrow(Rts_epiestim),
                           Rts_epiestim)


layout(matrix(1:2, nrow = 1))
P1 <- diag_plot(I_NB, logged = TRUE, max_x=1, dist = 'poisson', Rt = Rts_epiestim)
P2 <- diag_plot(I_NB, logged = FALSE, max_x=.1, dist = 'poisson', Rt = Rts_epiestim)

P1
P2


P1$check$v1 <- P1$check$residual^2/P1$check$Exp

plot((P1$check$Rt),(P1$check$v1))

m1<-lm(P1$check$v1~1+P1$check$Rt )
summary(m1)
m0<-lm(P1$check$v1~1  )
summary(m0)

a1 <- anova(m0,m1, test = 'Chisq')
a1
if(a1$`Pr(>Chi)`[2]<0.05){
  print(paste0('likely overdispersion, guestimate of dipersion parameter: ',round(1/m1$coefficients[2],digits = 2)) )
}else{
  print('unlikely overdispersion')
}


```


## Run MCMCEpiEstim 

```{r}

rep <- 1e3

res_MCMC_EpiEstim_Poiss1 <- fct_MCMC_EpiEstim(I0 = sim$I0 , I = sim$res$I_obs, #I_true ,
                                              t_window = t_window,
                                             mean_prior = mean_prior, std_prior = std_prior,
                                             res_EpiEstim = res_EpiEstim, overdispersion = FALSE, 
                                             rep = rep, thin = 10, param_agg = FALSE)

res_MCMC_EpiEstim_Poiss2 <- fct_MCMC_EpiEstim(I0 = sim$I0 , I = sim$res$I_obs, #I_true , 
                                              t_window = t_window,
                                             mean_prior = mean_prior, std_prior = std_prior,
                                             res_EpiEstim = res_EpiEstim, overdispersion = FALSE, 
                                             rep = rep, thin = 10, param_agg = TRUE)

```


```{r}

res_MCMC_EpiEstim_NB1 <- fct_MCMC_EpiEstim(I0 = sim$I0 , I = sim$res$I_obs,#I_true  ,
                                           t_window = t_window,
                                          mean_prior = mean_prior, std_prior = std_prior,
                                          res_EpiEstim = res_EpiEstim, overdispersion = TRUE,
                                          rep = rep, thin = 10, param_agg = FALSE)


res_MCMC_EpiEstim_NB2 <- fct_MCMC_EpiEstim(I0 = sim$I0 , I = sim$res$I_obs, #I_true  ,
                                           t_window = t_window,
                                          mean_prior = mean_prior, std_prior = std_prior,
                                          res_EpiEstim = res_EpiEstim, overdispersion = TRUE,
                                          rep = rep, thin = 10, param_agg = TRUE)


```


```{r}

dics <- data.frame(model = c('Pois1','Pois2','NB1','NB2'),
                   DIC = c(res_MCMC_EpiEstim_Poiss1$DIC[1],
                           res_MCMC_EpiEstim_Poiss2$DIC[1],
                           res_MCMC_EpiEstim_NB1$DIC[1],
                           res_MCMC_EpiEstim_NB2$DIC[1]))
dics
```


## check

```{r}
LogL <- cbind(rowSums(res_MCMC_EpiEstim_Poiss1$logL),
              rowSums(res_MCMC_EpiEstim_Poiss2$logL),
              rowSums(res_MCMC_EpiEstim_NB1$logL),
              rowSums(res_MCMC_EpiEstim_NB2$logL))
R1 <- cbind(res_MCMC_EpiEstim_Poiss1$theta_R[,1],
            res_MCMC_EpiEstim_Poiss2$theta_R[,1],
            res_MCMC_EpiEstim_NB1$theta_R[,1],
            res_MCMC_EpiEstim_NB2$theta_R[,1])

layout(matrix(1:4,2,2))
plot(LogL[,1],ylim = range(c(LogL)))
lines(LogL[,2],col = 'red3',type = 'p',pch = 12)
lines(LogL[,3],col = 'blue3',type = 'p',pch = 12)
lines(LogL[,4],col = 'green3',type = 'p',pch = 12)

plot(R1[,1],ylim = range(c(R1)))
lines(R1[,2],col = 'red3',type = 'p',pch = 12)
lines(R1[,3],col = 'blue3',type = 'p',pch = 12)
lines(R1[,4],col = 'green3',type = 'p',pch = 12)


# others
acf(res_MCMC_EpiEstim_Poiss1$theta_R[,1])
acf(res_MCMC_EpiEstim_NB1$theta_R[,1])

plot(res_MCMC_EpiEstim_NB1$theta_over , 
     ylim = range(c(res_MCMC_EpiEstim_NB1$theta_over,
                    res_MCMC_EpiEstim_NB2$theta_over)))

lines(res_MCMC_EpiEstim_NB2$theta_over,col = 'red3',type = 'p',pch = 12)

quantile(res_MCMC_EpiEstim_NB1$theta_over, c(.5,.025,.975))
quantile(res_MCMC_EpiEstim_NB2$theta_over, c(.5,.025,.975))
```

```{r}
# function for plots compare results
t_max <- length(res_EpiEstim[[1]]$t)
n_location <- sim$input$n_locations
  
comp_plot <- function(f1,f2){
  
  if(f2 %in% c("Quantile.0.025(R)","Quantile.0.975(R)")){
    if(f2 %in%"Quantile.0.025(R)"){
      Rts_mcmc_pois <- apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,quantile,0.025,na.rm = TRUE)
      Rts_mcmc_NB <- apply(res_MCMC_EpiEstim_NB1$theta_R,2,quantile,0.025,na.rm = TRUE)
    }else{
      Rts_mcmc_pois <- apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,quantile,0.975,na.rm = TRUE)
      Rts_mcmc_NB <- apply(res_MCMC_EpiEstim_NB1$theta_R,2,quantile,0.975,na.rm = TRUE)
    }
  }else{
    Rts_mcmc_pois <- apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,f1)
    Rts_mcmc_NB <- apply(res_MCMC_EpiEstim_NB1$theta_R,2,f1)
  }
  Rts_epiestim <- matrix(unlist(lapply(res_EpiEstim, "[", ,f2)), nrow = t_max, ncol = n_location, byrow = FALSE)
  Rts_epiestim <- c(Rts_epiestim[(t_window+I0$timespan):nrow(Rts_epiestim),])
  
  if(length(Rts_epiestim) == n_location*length(Rts_mcmc_NB)){
    Rts_mcmc_pois <- rep(Rts_mcmc_pois,n_location)
    Rts_mcmc_NB <- rep(Rts_mcmc_NB,n_location)
  }
  
  plot(log10(Rts_epiestim), log10(Rts_mcmc_pois), bty = 'n', main = f2, col = 'blue3',pch=3)
  lines(log10(Rts_epiestim), log10(Rts_mcmc_NB), type = 'p', col = 'red3',pch=4)
  abline(a = c(0,1), col = 'grey50')
  # range((Rts_epiestim-Rts_mcmc), na.rm = TRUE)
  
}


layout(matrix(1:9,3,3, byrow = TRUE))
comp_plot(f1 = mean,f2 = 'Mean(R)')
comp_plot(f1 = median,f2 = "Median(R)")
comp_plot(f1 = sd,f2 = "Std(R)")
comp_plot(f1 = mean,f2 = "Quantile.0.025(R)")
comp_plot(f1 = mean,f2 = "Quantile.0.975(R)")

legend('topleft',c('Poiss','NB'),pch = c(3,4), col = c('blue3','red3'),bty = 'n')

```



```{r}
layout(matrix(1:4,2,2))
# time windows
  t_start <- seq(sim$I0$timespan+1, t_max-t_window+1,by = 1)        
  t_end <- t_start + t_window - 1     
  n_tw <- length(t_start)

# sim vs  epiestim
Rts_epiestim_med <- matrix(unlist(lapply(res_EpiEstim, "[", ,"Median(R)")), 
                           nrow = t_max, ncol = n_location, byrow = FALSE)
Rts_epiestim_mean <- matrix(unlist(lapply(res_EpiEstim, "[", ,"Mean(R)")), 
                            nrow = t_max, ncol = n_location, byrow = FALSE)

plot(rep(seq(1,t_max),n_location),c(Rts_epiestim_med), 
     main = 'median/mean EpiEstim vs sim',bty = 'n',pch=3,col = 'blue3')
lines(rep(seq(1,t_max),n_location),c(Rts_epiestim_mean),type = 'p',pch=4, col = 'red3')
lines(Rt$t, Rt$Rt, col = 'grey50')

# sim vs MCMC Poiss
Rts_mcmc_poiss_med <- matrix(apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,median), 
                             nrow =  n_tw, ncol = n_location, byrow = FALSE)
Rts_mcmc_poiss_mean <- matrix(apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,mean), 
                              nrow =  n_tw, ncol = n_location, byrow = FALSE)

plot(rep(t_end,n_location),c(Rts_mcmc_poiss_med), 
     main = 'median/mean MCMC Pois vs sim',bty = 'n',pch=3,col = 'blue3')
lines(rep(t_end,n_location),c(Rts_mcmc_poiss_mean),type = 'p',pch=4, col = 'red3')
lines(Rt$t, Rt$Rt, col = 'grey50')

# sim vs MCMC NB
Rts_mcmc_NB_med <- matrix(apply(res_MCMC_EpiEstim_NB1$theta_R,2,median), 
                             nrow =  n_tw, ncol = n_location, byrow = FALSE)
Rts_mcmc_NB_mean <- matrix(apply(res_MCMC_EpiEstim_NB1$theta_R,2,mean), 
                              nrow =  n_tw, ncol = n_location, byrow = FALSE)

plot(rep(t_end,n_location),c(Rts_mcmc_NB_med), 
     main = 'median/mean MCMC NB vs sim',bty = 'n',,pch=3,col = 'blue3')
lines(rep(t_end,n_location),c(Rts_mcmc_NB_mean),type = 'p',pch=4, col = 'red3')
lines(Rt$t, Rt$Rt, col = 'grey50')

legend('topleft',c('median','mean'),pch = c(3,4), col = c('blue3','red3'),bty = 'n')

```


```{r}
I_NB <- as.matrix(sim$res$I_true)

Rts_P1 <- matrix(apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,median),
                                   nrow =  n_tw, ncol = n_location, byrow = FALSE)

Rts_P1 <- Rts_P1[seq(1,nrow(Rts_P1),by = 7),]
Rts_P1 <- Rts_P1[rep(seq(1:nrow(Rts_P1)),each = 7),]
Rts_P1 <- data.frame(t = 1:nrow(Rts_P1),
                     Rts_P1)


layout(matrix(1:2, nrow = 1))
P1 <- diag_plot(I_NB, logged = TRUE, max_x=1, dist = 'poisson', Rt = Rts_P1)
P2 <- diag_plot(I_NB, logged = FALSE, max_x=.1, dist = 'poisson', Rt = Rts_P1)

P1
P2

```


```{r}

Rts_nb1 <- matrix(apply(res_MCMC_EpiEstim_NB1$theta_R,2,median),
                                   nrow =  n_tw, ncol = n_location, byrow = FALSE)

Rts_nb1 <- Rts_nb1[seq(1,nrow(Rts_nb1),by = 7),]
Rts_nb1 <- Rts_nb1[rep(seq(1:nrow(Rts_nb1)),each = 7),]
Rts_nb1 <- data.frame(t = 1:nrow(Rts_nb1),
                     Rts_nb1)

layout(matrix(1:2, nrow = 1))
NB1 <- diag_plot(I_NB, logged = TRUE, max_x=1, dist = 'nb', Rt = Rts_nb1,k = median(res_MCMC_EpiEstim_NB1$theta_over0))
NB2 <- diag_plot(I_NB, logged = FALSE, max_x=.1, dist = 'nb', Rt = Rts_nb1, k = median(res_MCMC_EpiEstim_NB1$theta_over0))

NB1
NB2

```





# check MCMC with Poisson simulation

# simulate

```{r}

Scenarios <- as.data.frame(readxl::read_xlsx('../overdispersion-epidemic-data/overdispersion_scenarios.xlsx',
                               sheet = 1, range = 'A1:I60'))
si <- c(0,1)

i=57

# print(i)
input <- Scenarios[i,]

input$R_min <- 0.65
input$k <- 1e3

I0 <- incidence::as.incidence(x = input$I0, dates = 1, interval = 1)

Rt <- MCMCEpiEstim::Rt_linear(inp = input)

res <- MCMCEpiEstim::project_fct(I0 = I0, Rt = Rt, 
                                 n_loc = input$n_locations,
                                 t_max = nrow(Rt), si = si,p = input$pi,
                                 model = 'negbin',over = input$k)
sim <- list(input = input, I0 = I0, Rt = Rt, res = res)


matplot((sim$res$I_true[,-1]))
matplot((sim$res$I_obs[,-1]))
input
  
```


```{r}


# parameters
# for EpiEstim
t_window <- 7
mean_prior <- 5
std_prior <- 5

res_EpiEstim <- fct_EpiEstim(I0 = sim$I0 , I = sim$res$I_true , t_window = t_window, 
                             mean_prior = mean_prior, std_prior = std_prior,
                             si = si, overlap = FALSE)
aa <- sim$res$I_true
ab <- res_EpiEstim[[1]]

plot(Rt$t,Rt$Rt,type = 'l', ylim = c(0,2.5))

for(i in 1:sim$input$n_locations){
  
  Hmisc::errbar(x = res_EpiEstim[[i]]$t_start, 
                y = res_EpiEstim[[i]]$`Median(R)`,
                yplus = res_EpiEstim[[i]]$`Quantile.0.975(R)`,
                yminus = res_EpiEstim[[i]]$`Quantile.0.025(R)`,
                col = i, errbar.col = i, 
                add=TRUE)
}



```

diagnoistics for epiEstim

```{r}

I_NB <- as.matrix(sim$res$I_true)

Rts_epiestim <- matrix(unlist(lapply(res_EpiEstim, "[", ,'Mean(R)')), 
                       nrow = t_max, ncol = n_location, byrow = FALSE)

Rts_epiestim <- Rts_epiestim[seq(1,nrow(Rts_epiestim),by = 7),]
Rts_epiestim <- Rts_epiestim[rep(seq(2,nrow(Rts_epiestim)),each = 7),]

Rts_epiestim <- data.frame(t = 1:nrow(Rts_epiestim),
                           Rts_epiestim)


layout(matrix(1:2, nrow = 1))
P1 <- diag_plot(I_NB, logged = TRUE, max_x=1, dist = 'poisson', Rt = Rts_epiestim)
P2 <- diag_plot(I_NB, logged = FALSE, max_x=.1, dist = 'poisson', Rt = Rts_epiestim)

P1$res_table
P2$res_table


P1$check$v1 <- P1$check$residual^2/P1$check$Exp

plot(P1$check$Rt,P1$check$v1)

m1<-lm(P1$check$v1~1+P1$check$Rt )
summary(m1)
m0<-lm(P1$check$v1~1  )
summary(m0)

a1 <- anova(m0,m1, test = 'Chisq')
a1
if(a1$`Pr(>Chi)`[2]<0.05){
  print(paste0('likely overdispersion, guestimate of dipersion parameter: ',round(1/m1$coefficients[2],digits = 2)) )
}else{
  print('unlikely overdispersion')
}

```


## Run MCMCEpiEstim 

```{r}

rep <- 1e3

res_MCMC_EpiEstim_Poiss1 <- fct_MCMC_EpiEstim(I0 = sim$I0 , I = sim$res$I_true , t_window = t_window,
                                             mean_prior = mean_prior, std_prior = std_prior,
                                             res_EpiEstim = res_EpiEstim, overdispersion = FALSE, 
                                             rep = rep, thin = 10, param_agg = FALSE)

res_MCMC_EpiEstim_Poiss2 <- fct_MCMC_EpiEstim(I0 = sim$I0 , I = sim$res$I_true , t_window = t_window,
                                             mean_prior = mean_prior, std_prior = std_prior,
                                             res_EpiEstim = res_EpiEstim, overdispersion = FALSE, 
                                             rep = rep, thin = 10, param_agg = TRUE)

```


```{r}

res_MCMC_EpiEstim_NB1 <- fct_MCMC_EpiEstim(I0 = sim$I0 , I = sim$res$I_true  , t_window = t_window,
                                          mean_prior = mean_prior, std_prior = std_prior,
                                          res_EpiEstim = res_EpiEstim, overdispersion = TRUE,
                                          rep = rep, thin = 10, param_agg = FALSE)


res_MCMC_EpiEstim_NB2 <- fct_MCMC_EpiEstim(I0 = sim$I0 , I = sim$res$I_true  , t_window = t_window,
                                          mean_prior = mean_prior, std_prior = std_prior,
                                          res_EpiEstim = res_EpiEstim, overdispersion = TRUE,
                                          rep = rep, thin = 10, param_agg = TRUE)


```


```{r}

dics <- data.frame(model = c('Pois1','Pois2','NB1','NB2'),
                   DIC = c(res_MCMC_EpiEstim_Poiss1$DIC[1],
                           res_MCMC_EpiEstim_Poiss2$DIC[1],
                           res_MCMC_EpiEstim_NB1$DIC[1],
                           res_MCMC_EpiEstim_NB2$DIC[1]))
dics
```


## check

```{r}
LogL <- cbind(rowSums(res_MCMC_EpiEstim_Poiss1$logL),
              rowSums(res_MCMC_EpiEstim_Poiss2$logL),
              rowSums(res_MCMC_EpiEstim_NB1$logL),
              rowSums(res_MCMC_EpiEstim_NB2$logL))
R1 <- cbind(res_MCMC_EpiEstim_Poiss1$theta_R[,1],
            res_MCMC_EpiEstim_Poiss2$theta_R[,1],
            res_MCMC_EpiEstim_NB1$theta_R[,1],
            res_MCMC_EpiEstim_NB2$theta_R[,1])

layout(matrix(1:4,2,2))
plot(LogL[,1],ylim = range(c(LogL)))
lines(LogL[,2],col = 'red3',type = 'p',pch = 12)
lines(LogL[,3],col = 'blue3',type = 'p',pch = 12)
lines(LogL[,4],col = 'green3',type = 'p',pch = 12)

plot(R1[,1],ylim = range(c(R1)))
lines(R1[,2],col = 'red3',type = 'p',pch = 12)
lines(R1[,3],col = 'blue3',type = 'p',pch = 12)
lines(R1[,4],col = 'green3',type = 'p',pch = 12)


# others
acf(res_MCMC_EpiEstim_Poiss1$theta_R[,1])
acf(res_MCMC_EpiEstim_NB1$theta_R[,1])

plot(res_MCMC_EpiEstim_NB1$theta_over , 
     ylim = range(c(res_MCMC_EpiEstim_NB1$theta_over,
                    res_MCMC_EpiEstim_NB2$theta_over)))

lines(res_MCMC_EpiEstim_NB2$theta_over,col = 'red3',type = 'p',pch = 12)

quantile(res_MCMC_EpiEstim_NB1$theta_over, c(.5,.025,.975))
quantile(res_MCMC_EpiEstim_NB2$theta_over, c(.5,.025,.975))
```

```{r}
# function for plots compare results
t_max <- length(res_EpiEstim[[1]]$t)
n_location <- sim$input$n_locations
  
comp_plot <- function(f1,f2){
  
  if(f2 %in% c("Quantile.0.025(R)","Quantile.0.975(R)")){
    if(f2 %in%"Quantile.0.025(R)"){
      Rts_mcmc_pois <- apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,quantile,0.025,na.rm = TRUE)
      Rts_mcmc_NB <- apply(res_MCMC_EpiEstim_NB1$theta_R,2,quantile,0.025,na.rm = TRUE)
    }else{
      Rts_mcmc_pois <- apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,quantile,0.975,na.rm = TRUE)
      Rts_mcmc_NB <- apply(res_MCMC_EpiEstim_NB1$theta_R,2,quantile,0.975,na.rm = TRUE)
    }
  }else{
    Rts_mcmc_pois <- apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,f1)
    Rts_mcmc_NB <- apply(res_MCMC_EpiEstim_NB1$theta_R,2,f1)
  }
  Rts_epiestim <- matrix(unlist(lapply(res_EpiEstim, "[", ,f2)), nrow = t_max, ncol = n_location, byrow = FALSE)
  Rts_epiestim <- c(Rts_epiestim[(t_window+I0$timespan):nrow(Rts_epiestim),])
  
  if(length(Rts_epiestim) == n_location*length(Rts_mcmc_NB)){
    Rts_mcmc_pois <- rep(Rts_mcmc_pois,n_location)
    Rts_mcmc_NB <- rep(Rts_mcmc_NB,n_location)
  }
  
  plot(log10(Rts_epiestim), log10(Rts_mcmc_pois), bty = 'n', main = f2, col = 'blue3',pch=3)
  lines(log10(Rts_epiestim), log10(Rts_mcmc_NB), type = 'p', col = 'red3',pch=4)
  abline(a = c(0,1), col = 'grey50')
  # range((Rts_epiestim-Rts_mcmc), na.rm = TRUE)
  
}


layout(matrix(1:9,3,3, byrow = TRUE))
comp_plot(f1 = mean,f2 = 'Mean(R)')
comp_plot(f1 = median,f2 = "Median(R)")
comp_plot(f1 = sd,f2 = "Std(R)")
comp_plot(f1 = mean,f2 = "Quantile.0.025(R)")
comp_plot(f1 = mean,f2 = "Quantile.0.975(R)")

legend('topleft',c('Poiss','NB'),pch = c(3,4), col = c('blue3','red3'),bty = 'n')

```



```{r}
layout(matrix(1:4,2,2))
# time windows
  t_start <- seq(sim$I0$timespan+1, t_max-t_window+1,by = 1)        
  t_end <- t_start + t_window - 1     
  n_tw <- length(t_start)

# sim vs  epiestim
Rts_epiestim_med <- matrix(unlist(lapply(res_EpiEstim, "[", ,"Median(R)")), 
                           nrow = t_max, ncol = n_location, byrow = FALSE)
Rts_epiestim_mean <- matrix(unlist(lapply(res_EpiEstim, "[", ,"Mean(R)")), 
                            nrow = t_max, ncol = n_location, byrow = FALSE)

plot(rep(seq(1,t_max),n_location),c(Rts_epiestim_med), 
     main = 'median/mean EpiEstim vs sim',bty = 'n',pch=3,col = 'blue3')
lines(rep(seq(1,t_max),n_location),c(Rts_epiestim_mean),type = 'p',pch=4, col = 'red3')
lines(Rt$t, Rt$Rt, col = 'grey50')

# sim vs MCMC Poiss
Rts_mcmc_poiss_med <- matrix(apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,median), 
                             nrow =  n_tw, ncol = n_location, byrow = FALSE)
Rts_mcmc_poiss_mean <- matrix(apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,mean), 
                              nrow =  n_tw, ncol = n_location, byrow = FALSE)

plot(rep(t_end,n_location),c(Rts_mcmc_poiss_med), 
     main = 'median/mean MCMC Pois vs sim',bty = 'n',pch=3,col = 'blue3')
lines(rep(t_end,n_location),c(Rts_mcmc_poiss_mean),type = 'p',pch=4, col = 'red3')
lines(Rt$t, Rt$Rt, col = 'grey50')

# sim vs MCMC NB
Rts_mcmc_NB_med <- matrix(apply(res_MCMC_EpiEstim_NB1$theta_R,2,median), 
                             nrow =  n_tw, ncol = n_location, byrow = FALSE)
Rts_mcmc_NB_mean <- matrix(apply(res_MCMC_EpiEstim_NB1$theta_R,2,mean), 
                              nrow =  n_tw, ncol = n_location, byrow = FALSE)

plot(rep(t_end,n_location),c(Rts_mcmc_NB_med), 
     main = 'median/mean MCMC NB vs sim',bty = 'n',,pch=3,col = 'blue3')
lines(rep(t_end,n_location),c(Rts_mcmc_NB_mean),type = 'p',pch=4, col = 'red3')
lines(Rt$t, Rt$Rt, col = 'grey50')

legend('topleft',c('median','mean'),pch = c(3,4), col = c('blue3','red3'),bty = 'n')

```


```{r}
I_NB <- as.matrix(sim$res$I_true)

Rts_P1 <- matrix(apply(res_MCMC_EpiEstim_Poiss1$theta_R,2,median),
                                   nrow =  n_tw, ncol = n_location, byrow = FALSE)

Rts_P1 <- Rts_P1[seq(1,nrow(Rts_P1),by = 7),]
Rts_P1 <- Rts_P1[rep(seq(1:nrow(Rts_P1)),each = 7),]
Rts_P1 <- data.frame(t = 1:nrow(Rts_P1),
                     Rts_P1)


layout(matrix(1:2, nrow = 1))
P1 <- diag_plot(I_NB, logged = TRUE, max_x=1, dist = 'poisson', Rt = Rts_P1)
P2 <- diag_plot(I_NB, logged = FALSE, max_x=.1, dist = 'poisson', Rt = Rts_P1)

P1$res_table
P2$res_table


```


```{r}

Rts_nb1 <- matrix(apply(res_MCMC_EpiEstim_NB1$theta_R,2,median),
                                   nrow =  n_tw, ncol = n_location, byrow = FALSE)

Rts_nb1 <- Rts_nb1[seq(1,nrow(Rts_nb1),by = 7),]
Rts_nb1 <- Rts_nb1[rep(seq(1:nrow(Rts_nb1)),each = 7),]
Rts_nb1 <- data.frame(t = 1:nrow(Rts_nb1),
                     Rts_nb1)

layout(matrix(1:2, nrow = 1))
NB1 <- diag_plot(I_NB, logged = TRUE, max_x=1, dist = 'nb', Rt = Rts_nb1,k = median(res_MCMC_EpiEstim_NB1$theta_over0))
NB2 <- diag_plot(I_NB, logged = FALSE, max_x=.1, dist = 'nb', Rt = Rts_nb1, k = median(res_MCMC_EpiEstim_NB1$theta_over0))

NB1$res_table
NB2$res_table

```
