---
title: "MCMCEpiEstim: package overview"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette:
     toc: true
     toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*MCMCEpiEstim* allows:

- inferring Rt on multiple locations using EpiEstim independelty on each location.

- inferring Rt on multiple locations using an MCMC implementation of EpiEstim and allowing Rt variations 
to be the same/different across locations. The implementation can account for a given level reporting which may change over time. 

- jointly inferring Rt and the level of superspreading on multiple locations using an MCMC implementation of EpiEstim and allowing Rt variations to be the same/different across locations. The implementation can account for a given level reporting which may change over time.

In the 3 implementations:

- the serial interval is assumed to be the non-parametric version of EpiEstim,

- the time window can be sliding overlapping or non-overlapping (the latter returning estimates of Rt at the end of each time window),

- level of superspreading is always assumed constant over time and across locations.

Below, we provide a typical workflow for using the package

# Installing the package

To install the current stable, CRAN version of the package, type:
```{r install, eval=FALSE}
install.packages("MCMCEpiEstim")
```

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:
```{r install2, eval=FALSE}
devtools::install_github("pnouvellet/MCMCEpiEstim",build = TRUE, force = TRUE)

```

Note that this requires the package *devtools* installed.

# Simulating incidence data

This section illustrates how to simulate incidence data for a time-varying Rt, with a given level of reporting and superspreading.

```{r}
library(MCMCEpiEstim)

# a serial interal
si <- c(0,1)
# simualte an Rt
input <- data.frame(R_min = 0.95,
                    R_max = 1.1,
                    k = .1,
                    period = 24, 
                    step = 4,
                    I0 = 100,
                    n_loc = 10,
                    pi = 0.5)
Rt <- Rt_linear(inp = input, n_week = input$period)
plot(Rt$t, Rt$Rt,type='l', xlab = 'time',ylab = 'Rt')
# set initial incidence
I0 <- incidence::as.incidence(x = input$I0, dates = 1, interval = 1)

# simulate for 10 locations
res <- MCMCEpiEstim::project_fct(I0 = I0,
                                 Rt = Rt,
                                 n_loc = input$n_loc,
                                 t_max = nrow(Rt),
                                 si = si,
                                 p = input$pi,
                                 model = 'negbin',
                                 over = input$k, 
                                 k.seed = 1)

matplot((res$I_true[,-1]), xlab = '', ylab = 'cases', main = 'true cases')
matplot((res$I_obs[,-1]), xlab = '', ylab = 'cases', main = 'observed cases')

```


# Inferring Rt using the multi-location EpiEstim

```{r}

# parameters for EpiEstim
t_window <- 7
mean_prior <- 5
std_prior <- 5

res_EpiEstim <- fct_EpiEstim(I0_t_import = length(input$I0) , 
                             I = res$I_obs, 
                             t_window = t_window, 
                             mean_prior = mean_prior, 
                             std_prior = std_prior,
                             si = si, 
                             overlap = FALSE)


plot(Rt$t,Rt$Rt,type = 'l', ylim = c(0,2.5), xlab = 'time',ylab = 'Rt')

for(i in 1:input$n_loc){
  
  Rt_EpiEstim <- res_EpiEstim[[i]]
  # remove espitmate when overall infectivity was 0 (would return only prior distribution of Rt)
  f <- (which(Rt_EpiEstim$Oi == 0))
  Rt_EpiEstim[f,c("Mean(R)","Std(R)","Quantile.0.025(R)","Median(R)","Quantile.0.975(R)")] <- NA
  
  Hmisc::errbar(x = Rt_EpiEstim$t_start, 
                y = Rt_EpiEstim$`Median(R)`,
                yplus = Rt_EpiEstim$`Quantile.0.975(R)`,
                yminus = Rt_EpiEstim$`Quantile.0.025(R)`,
                col = i, errbar.col = i, 
                add=TRUE)
}



```
## Diagnostics for MCMCEpiEstim

```{r}

D1 <- diag_plot(res = res_EpiEstim,logged = FALSE,max_x = 1,dist = 'poisson')
D1$res_table

D1_logged <- diag_plot(res = res_EpiEstim,logged = TRUE,max_x = 1,dist = 'poisson')
D1_logged$res_table

```


# Jointly inferring Rt and the level of superspreading 


## first wrongy assuming 100% reporting

```{r}

rep <- 1e4 # number of iteration for the posterior distribution

# inferrence of Rt and overdispersion wrongly assuming 100% reporting
res_MCMC_EpiEstim_NB1 <- fct_MCMC_EpiEstim(I0_t_import = length(input$I0) ,
                                           I = res$I_obs,
                                           t_window = t_window,
                                           mean_prior = mean_prior, std_prior = std_prior,
                                           res_EpiEstim = res_EpiEstim, 
                                           overdispersion = TRUE,
                                           rep = rep, thin = 10, 
                                           param_agg = TRUE, 
                                           p_reps = 1,
                                           overlap = FALSE)



```
```{r}
k <- median(res_MCMC_EpiEstim_NB1$theta_over_thinned)
quantile(res_MCMC_EpiEstim_NB1$theta_over_thinned,c(.5,.025,.975))

D2 <- diag_plot(res = res_EpiEstim,logged = FALSE,max_x = 1,
                dist = 'nb', k = k, res_MCMC = res_MCMC_EpiEstim_NB1)
D2$res_table

D2_logged <- diag_plot(res = res_EpiEstim,logged = TRUE,max_x = 1,
                       dist = 'nb', k = k, res_MCMC = res_MCMC_EpiEstim_NB1)
D2_logged$res_table

```


## Then proper reporting assumption

```{r}
res_MCMC_EpiEstim_NB2 <- fct_MCMC_EpiEstim(I0_t_import = length(input$I0) ,
                                           I = res$I_obs, 
                                           t_window = t_window,
                                           mean_prior = mean_prior, std_prior = std_prior,
                                           res_EpiEstim = res_EpiEstim, 
                                           overdispersion = TRUE,
                                           rep = rep, thin = 10, 
                                           param_agg = TRUE,
                                           p_reps = input$pi,
                                           overlap = FALSE)



```


```{r}
k <- median(res_MCMC_EpiEstim_NB2$theta_over_thinned)
quantile(res_MCMC_EpiEstim_NB2$theta_over_thinned,c(.5,.025,.975))

D3 <- diag_plot(res = res_EpiEstim,logged = FALSE,max_x = 1,
                dist = 'nb', k = k, res_MCMC = res_MCMC_EpiEstim_NB2)
D3$res_table

D3_logged <- diag_plot(res = res_EpiEstim,logged = TRUE,max_x = 1,
                       dist = 'nb', k = k, res_MCMC = res_MCMC_EpiEstim_NB2)
D3_logged$res_table

```



# real world application - Ebola


