---
title: "MCMCEpiEstim: package overview"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette:
     toc: true
     toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*MCMCEpiEstim* allows:

- inferring Rt on multiple locations using EpiEstim independelty on each location.

- inferring Rt on multiple locations using an MCMC implementation of EpiEstim and allowing Rt variations 
to be the same/different across locations. The implementation can account for a given level reporting which may change over time. 

- jointly inferring Rt and the level of superspreading on multiple locations using an MCMC implementation of EpiEstim and allowing Rt variations to be the same/different across locations. The implementation can account for a given level reporting which may change over time.

In the 3 implementations:

- the serial interval is assumed to be the non-parametric version of EpiEstim,

- the time window can be sliding overlapping or non-overlapping (the latter returning estimates of Rt at the end of each time window),

- level of superspreading is always assumed constant over time and across locations.


# Installing the package

To install the current stable, CRAN version of the package, type:
```{r install, eval=FALSE}
install.packages("MCMCEpiEstim")
```

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:
```{r install2, eval=FALSE}
devtools::install_github("reconhub/MCMCEpiEstim")
```

Note that this requires the package *devtools* installed.

# Main visible functions of the package

The main functions of the package include:

...

Also consider adding some examples.

# Main internal functions

In case you want to describe internal function as well.
... 


